---
title: "Basic Markov Chain model"
author: "LZ"
date: "16 4 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Basic idea of Markov Chain model

The model assumes that the actions are memoryless, i.e., for a given state of 
the model any action that has happened before has no influence on what will happen
next. 
For example, when working with the states `(Ball in) Midfield`, `(Ball in) Box`, `(Ball on) Wing`, 
`Goal`, `Ball lost`, the probabilities of moving from one state to another could 
be described by the following *transition* matrix

```{r cars}
D <- matrix(c(0.25, 0.2, 0.1, 0.05, 0.4, 0.1, 0.25, 0.2, 0.15, 0.3, 0.1, 0.1, 0.25,
              0.05, 0.5, 0,0,0,1,0,0,0,0,0,1), byrow = TRUE, ncol = 5)
rownames(D) <- colnames(D) <- c("M", "B", "W", "G", "L")
D
```

The states `Goal` and `Ball Lost` are absorbing states.
When we want to find out the *expected threat* of having the ball within a certain state, 
this translates to finding the solution of the equation explained below. 
Let $\mathbf{x}_T = (x_M, x_B, x_W)^\top$ denote the vector of expected threat with entries for 
expected threat in Midfield, in Box and on Wing, respectively. 
When the ball is in midfield, the threat of having the ball at that position is the 
probability of scoring, plus the probability of passing/carrying the ball within the midfield
area times the threat of having the ball in the midfield are, plus the probability of 
passing/carrying the ball to the box times
the threat of having the ball in the box, plus the probability of passing/carrying the
ball to the wing times the threat of having the ball at the wing. Similar for having the 
ball at the wing or in the box. Mathematically, this can be expressed in the 
following equations:

\[ \begin{aligned}
x_M &= 0.05 + 0.25x_M 0.2x_B + 0.1x_W \\
   x_B &= 0.15 + 0.1x_M + 0.25 x_B + 0.2x_W\\
   x_W &= 0.05 + 0.1x_M + 0.1x_B + 0.25x_W.
   \end{aligned}
\]

In matrix notation, this can be written as

\[ \begin{aligned}
    \mathbf{x}_T  &= A\mathbf{x}_T  + g \\
    \Leftrightarrow \ (I - A)\mathbf{x}_T  &= g,
    \end{aligned}
\]

where 

```{r}
A <- D[1:3, 1:3]
A  
```
and 
$g = (0.05, 0.15, 0.05)^\top$.
This equation needs to be solved for $\mathbf{x}_T $.

```{r, echo=FALSE}
g <- c(0.05, 0.15, 0.05)
```

```{r}
solve(diag(1,3) - A, g)
```

### Solving iteratively

Instead of using linear algebra, one can solve the equation iteratively until convergence.

To do so, we initialise $\mathbf{x}_T^{(0)} = (0, 0, 0)^\top$ and run the equation 
iteratively, i.e. 

```{r}
xT <- c(0, 0, 0)
eps <- 1e-10
for(i in 1:100) {
  xT_old <- xT
  xT  <- A %*% xT + g
  if(sum((xT - xT_old)^2) < eps) { 
    break()
  }
}
i
colnames(xT) <- "xT"
t(xT)
```
