---
title: "Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE, message=FALSE}
library(soccerlytics)
library(tidyverse)
```
### Load data

The visualistions will be carried out based on the publicly availabe event data set 
from the 2012/2013 Champion's League Final. 
The data is made available by StatsBomb.

```{r, warning=FALSE, message=FALSE}
# load Data
FreeComp <- StatsBombR::FreeCompetitions()
FreeMatch <-  StatsBombR::FreeMatches(FreeComp %>% filter(season_name == "2012/2013"))

# filter match with id 18240 (CL final)
FreeMatch <- FreeMatch %>% filter(match_id == 18240)
eventData <- StatsBombR::get.matchFree(FreeMatch)

## clean eventData (unnest several columns)
eventData <- StatsBombR::allclean(eventData)

```

### Tactical line up

To plot the tactical line up, we first extract the lineup dataframes within the 
event dataset and use the preparation function `prep_formation_data` which can then 
be passed to the `plot_formation` function, along with the additional information 
of which player was substituted, here provided in the dataframe `subPlayers`.
```{r, fig.dim=c(12,8)}

dfForm <- prep_formation_data(eventData$tactics.lineup[[1]], team.name = "BVB") %>% 
  bind_rows(prep_formation_data(eventData$tactics.lineup[[2]], team.name = "Bayern Munich"))
subPlayers <- eventData %>% filter(type.name == "Substitution") %>% 
  select(team.name, player.name, minute, second, starts_with("Sub"))
plot_formation(dfFormation = dfForm, colourHome = "yellow", colourHome2 = "black",
               colourAway = "red", hometeam = "BVB", 
               awayteam = "Bayern Munich", subPlayers = subPlayers, 
               formHome = "4-2-3-1", formAway = "4-2-3-1")
```
